Minimax-optimal Laplacian Eigenmaps regression over Sobolev Spaces with Neighborhood Graphs

### 2021-03-18 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Finished editting higher-order section of Laplacian eigenmaps paper.

In this paper we focus on the theoretical statistical properties of \textcolor{red}{graph-based regression}. We assume that in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$. The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which takes the eigenvectors of the graph Laplacian $L$. The Laplacian $L$ is a difference operator, acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
\begin{equation}
\label{eqn:graph_laplacian}
Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)W_{ij}. 
\end{equation}
The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$ \textcolor{red}{(...)}. The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$. Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the corresponding eigenvector.

@@

\begin{table}
	\begin{center}
		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $4 <\dim(\mc{X}) < 4s$  & \textcolor{red}{(?)}& \textcolor{red}{(?)} \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & \textcolor{red}{(?)} \\
			& $4 < \dim(\mc{X}) < 4s$ & \textcolor{red}{(?)} & \textcolor{red}{(?)} \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:testing_rates}
\end{table}

\paragraph{Related Work.} 
\begin{itemize}
	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $u_k$, in the sense that \textcolor{red}{(...)}. \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$. This suggests that Laplacian eigenmaps may be good alternatives to classical spectral series methods when the latter cannot be implemented (e.g. because $P$ or $\mc{X}$ are unknown.) 
	\item There has been limited analysis of the estimator $\wh{f}$. \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity. In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$. \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but they derive suboptimal rates for the supervised problem. As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$. The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty along with a data-fidelity term,
	\begin{equation*}
	\wh{f}_{\mathrm{LS}} := \|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
	\end{equation*}
	\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$. In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$. 
\end{itemize}

@@

\begin{itemize}
	\item Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, its norm, and its normed ball.
	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
\end{itemize}

@@

Let us explain the reason we require boundary conditions in the flat Euclidean case when $s > 1$, by considering an alternative way to characterize Sobolev spaces using a relation to sequence spaces.\footnote{This is equivalent to characterizing Sobolev spaces using the Fourier transform, but we find it more transparent to instead work directly with sequence spaces.} Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the Laplace-Beltrami eigenvector equation (with Neumann boundary conditions), 
\begin{equation*}
\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$},
\end{equation*}
where $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$ to the boundary of $\mc{X}$.
We define an ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ as $\Theta := \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} \lambda_k(\Delta_P)^{2s} \theta_k^2 < \infty\}$. Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(P)$ which satisfy the following \emph{Neumann} boundary conditions,
\begin{equation}
\label{eqn:condition}
\frac{\partial^j}{\partial {\bf n}^j}f = 0 \quad \textrm{for $j = 1,3,5,\ldots,\floor{s/2}$.}
\end{equation}
As we have already reviewed, the eigenvectors $v_k$ of the graph Laplacian $L_{n,\varepsilon}$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$. As a result, is natural that they accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(P)$. Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that the even derivatives be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps. We will stick to the zero-trace condition, which greatly eases some steps in our proofs.

@@

The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral series estimator.  Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, a fact that we discuss in more detail later. A key consequence of this scaling is that the Sobolev ball $H_0^s(P) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$. By reducing the regression problem to an estimation problem in a Gaussian sequence model \textcolor{red}{(see e.g. Tsybakov,Johnstone,Gine)}, it can be shown the estimator
\begin{equation*}
\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
\end{equation*}
achieves the minimax rate over $\mc{H}^s(P;M)$, and the same upper bound on the minimax rate therefore applies to $H_0^s(P;M)$.

@@

@@

\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate, when error is measured in $\|\cdot\|_{n}$.
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{3}
	\item 
	\label{asmp:parameters_estimation_fo} 
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
	\begin{equation}\\
	\label{eqn:radius_fo} 
	\biggl(\frac{\log n}{n}\biggr)^{1/d} \ll \varepsilon \ll \min\{1,K^{-1/d}\},
	\end{equation}
	and 
	\begin{equation}
	\label{eqn:eigenvector_estimation_fo} 
	K \asymp \min\Bigl\{(M^2n)^{d/(2 + d)} \vee 1, n\Bigr\}
	\end{equation}
\end{enumerate}
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_fo}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$. There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_fo}
	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2/(2 + d)} \wedge 1\Bigr) \vee \frac{1}{n},
	\end{equation}
	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
\end{theorem}
From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, and is thus an in-sample minimax rate-optimal estimator.

Some other remarks:
\begin{itemize}
	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}. In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. However, in truth these are edge cases, \textcolor{red}{(...)}.
	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected. On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$. Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$. Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$. 
\end{itemize}

\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
\begin{equation*}
t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}
\end{equation*}
This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test. Moreover, when $\varepsilon$ and $K$ are chosen appropriately, the test $\varphi$ has negligible Type II error against alternatives separated from the null by at least $\|f_0\|_{P}^2 \gtrsim M^2(M^2n)^{-4/(4 + d)}$, whenever $d < 4$.

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{4}
	\item 
	\label{asmp:parameters_testing_fo}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_fo}
	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}
	\end{equation}
\end{enumerate}
In the following, let $p(d) = \min\{1,2/d\}$.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_fo}
	Fix $a,b \in (0,1)$. Suppose Model~\ref{def:model_flat_euclidean}. Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test. Suppose additionally $f_0 \in H^1(P,M)$, and that $d < 4$. Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
	\|f_0\|_n^2 \geq \frac{C}{\delta}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{1}{\delta n^{p(d)}} \biggr) \vee \frac{1}{n}
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the most important takeaway is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(\mc{X},M)$.

>>>>

In this paper we focus on the theoretical statistical properties of \textcolor{red}{graph-based regression}. We assume that in addition to the design points $X_1,\ldots,X_n$ one observes real-valued responses $Y_1,\ldots,Y_n$, and seeks to learn the unknown regression function $f_0(x) := E[Y|X = x]$. The methods we will study are based on \emph{Laplacian eigenmaps}, first introduced by \cite{belkin03a}, which takes the eigenvectors of the graph Laplacian $L$. The Laplacian $L$ is a difference operator, acting on functions $f: \{X_1,\ldots,X_n\} \to \Reals$ as follows,
\begin{equation}
\label{eqn:graph_laplacian}
Lf(X_i) = \sum_{j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)W_{ij}. 
\end{equation}
The Laplacian $L$ can be seen as a discretization of the Laplace-Beltrami operator $\Delta_P$, defined as
\begin{equation}
\label{eqn:fokker_planck_1}
\Delta_Pf= \frac{-1}{p} \mathrm{div}(p^2 \nabla f) = \Delta f - 2\frac{\nabla p^{\top} \nabla f}{p}.
\end{equation}
The eigenvectors of $L$ form an orthonormal basis of $L^2(X)$, and are approximations to the eigenvectors of $\Delta_P$. Their corresponding eigenvalues are estimates of~\textcolor{red}{(...)}, and provide a notion of smoothness for each eigenvector---the smaller the eigenvalue, the smoother the corresponding eigenvector.

@@

\begin{table}
	\begin{center}
		\begin{tabular}{p{.175\textwidth} p{.175\textwidth} | p{.14\textwidth} p{.12\textwidth} }
			Smoothness order & Dimension & Flat Euclidean (Model~\ref{def:model_flat_euclidean}) & Manifold (Model~\ref{def:model_manifold}) \\
			\hline
			\multirow{2}{*}{$s = 1$} & $\dim(\mc{X}) < 4$ & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $\dim(\mc{X}) \geq 4$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s = 2$ or $3$} & $\dim(\mc{X}) \leq 4$  & $\bm{n^{-4s/(4s + d)}}$ & $\bm{n^{-4s/(4s + m)}}$ \\
			& $4 <\dim(\mc{X}) < 4s$  & $n^{-2s/(2(s - 1) + d)}$ & $n^{-2s/(2(s - 1) + m)}$\\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
			\hline
			\multirow{3}{*}{$s > 3$} & $\dim(\mc{X}) \leq 4$ & $\bm{n^{-4s/(4s + d)}}$ & \textcolor{red}{(?)} \\
			& $4 < \dim(\mc{X}) < 4s$ & \textcolor{red}{(?)} & \textcolor{red}{(?)} \\
			& $\dim(\mc{X}) \geq 4s$ & $\bm{n^{-1/2}}$ & $\bm{n^{-1/2}}$ \\
		\end{tabular}
	\end{center}
	\caption{Summary of testing rates over Sobolev balls. Bold font marks minimax optimal rates. Rates when $d > 4s$ assume that $f_0 \in L^4(P,M)$. Although we suppress it for simplicity, in all cases the dependence of the error rate on the radius $M$ of the Sobolev ball is also optimal, as long as $M \gtrsim n^{-1/2}$.}
	\label{tbl:testing_rates}
\end{table}

\paragraph{Related Work.} 
\begin{itemize}
	\item \textcolor{red}{(TODO)} Summarize what is known on fixed graphs.
	\item \textcolor{red}{von Luxburg, Belkin, Slepcev} show that for a fixed index $k$, the eigenvector $v_k$ converges to the eigenfunction $\psi_k$, in the sense that \textcolor{red}{(...)}. \textcolor{red}{Shi, Burago, Garcia Trillos, Calder, Cheng} build on this, giving finite sample bounds, rates of convergence, and making statements uniform over $1 \leq k \leq k_{\max}$. This suggests that Laplacian eigenmaps may be good alternatives to classical spectral series methods when the latter cannot be implemented (e.g. because $P$ or $\mc{X}$ are unknown.) 
	\item There has been limited analysis of the estimator $\wh{f}$. \cite{zhou2011} consider a similar estimator, but in the semi-supervised setting where the number of unlabeled points grows to infinity. In this case the estimator reduces to a classical spectral series estimator, and there is no error incurred by approximating $\Delta_P$. \cite{lee2016} consider the \emph{diffusion maps} estimator---which uses the eigenvectors of a different normalization of the Laplacian $L$---in both the supervised and semi-supervised setups, but they derive suboptimal rates for the supervised problem. As far as we know, there has been no analysis of the test statistic $\wh{T}$ in the random design framework which we study.
	\item \textcolor{red}{(TODO)}: Alternatively, \cite{trillos2020} \textcolor{blue}{(Green 2021)} use the graph Laplacian operator to induce a penalty over functions $f: \{\bf X\} \to \Reals$. The \emph{Laplacian smoothing} estimator $\wh{f}_{\mathrm{LS}}$ is obtained by minimizing the sum of this penalty along with a data-fidelity term,
	\begin{equation*}
	\wh{f}_{\mathrm{LS}} := \|Y - f\|_n^2 + \lambda \dotp{L_{n,\varepsilon}f}{f}_n.
	\end{equation*}
	\textcolor{blue}{(Green 2021)} show that the resulting estimator is minimax optimal, but only for $s = 1$ and $d \leq 4$. In contrast, the Laplacian eigenmaps estimator is optimal for all $s$ and $d$. 
\end{itemize}

@@

\begin{itemize}
	\item Define $\Leb^2(\mc{X}), \Leb^2(P)$ and $\Leb^2(P_n)$. Define $C^k(\mc{X})$, its norm, and its normed ball.
	\item Define asymptotic notation: $\ll$, $\lesssim$ and $\asymp$.
	\item Define min and max notation: $\vee$ and $\wedge$.
\end{itemize}

@@

Let us explain the reason we require boundary conditions in the flat Euclidean case when $s > 1$, by considering an alternative way to characterize Sobolev spaces using a relation to sequence spaces.\footnote{This is equivalent to characterizing Sobolev spaces using the Fourier transform, but we find it more transparent to instead work directly with sequence spaces.} Let $(\lambda_1(\Delta_P),\psi_1),(\lambda_2(\Delta_P),\psi_2),\ldots$ be the solutions to the Laplace-Beltrami eigenvector equation (with Neumann boundary conditions), 
\begin{equation*}
\Delta_P\psi_k = \lambda_{k}(\Delta_P) \psi_k, \quad \frac{\partial}{\partial {\bf n}}\psi_k = 0~~\textrm{on $\partial \mc{X}$},
\end{equation*}
where $\partial/(\partial {\bf n})$ is the partial derivative operator in the direction of the normal vector $\mathbf{n}$ to the boundary of $\mc{X}$.
We define an ellipsoid $\Theta \subset \ell^2(\mathbb{N})$  using the eigenvalues $\lambda_1(\Delta_P),\lambda_2(\Delta_P),\ldots$ as $\Theta := \{\theta \in \ell^2(\mathbb{N}): \sum_{k = 1}^{\infty} \lambda_k(\Delta_P)^{2s} \theta_k^2 < \infty\}$. Then, the function class $\mc{H}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \theta \in \Theta\}$ consists precisely of those Sobolev functions $f \in H^s(P)$ which satisfy the following \emph{Neumann} boundary conditions,
\begin{equation}
\label{eqn:condition}
\frac{\partial^j}{\partial {\bf n}^j}f = 0 \quad \textrm{for $j = 1,3,5,\ldots,\floor{s/2}$.}
\end{equation}
As we have already reviewed, the eigenvectors $v_k$ of the graph Laplacian $L_{n,\varepsilon}$ converge to $\psi_k$ as $n \to \infty, \varepsilon \to 0$. As a result, is natural that they accurately approximate only those functions $f_0 \in \mc{H}^s(P) \subset H^s(P)$. Although the zero-trace boundary condition is more restrictive than~\eqref{eqn:condition}---since it also requires that derivatives of even-order be equal to $0$---the point is that some kind of boundary condition will be necessary in order to obtain optimal rates for spectral series methods, including Laplacian eigenmaps. We will stick to the zero-trace condition, since it greatly eases some of the steps in our proofs.

@@

The upper bound in~\eqref{eqn:sobolev_space_minimax_estimation_rate} can be certified by a particular spectral series estimator.  Under the assumptions of Model~\ref{def:model_flat_euclidean}, the eigenvalues $\lambda_k(\Delta_P)$ satisfy the Weyl's Law asymptotic scaling $\lambda_{k}(\Delta_P) \asymp k^{2/d}$, a fact that we discuss in more detail later. A key consequence of this scaling is that the Sobolev ball $H_0^s(P) \subset \mc{H}^s(P)$ continuously embeds into the Sobolev ellipsoid $\mc{\wt{H}}^s(P) = \{f = \sum_{k = 1}^{\infty} \theta_k \psi_k: \sum_{k = 1}^{\infty} (k - 1)^{2s/d} \theta_k^2\}$. By reducing the regression problem to an estimation problem in a Gaussian sequence model \textcolor{red}{(see e.g. Tsybakov,Johnstone,Gine)}, it can be shown the estimator
\begin{equation}
\label{eqn:spectral_series_estimator}
\wt{f} := \sum_{k = 1}^{K} \dotp{Y}{\psi_k}_n \psi_k,
\end{equation}
achieves the minimax rate over $\mc{H}^s(P;M)$, and the same upper bound on the minimax rate therefore applies to $H_0^s(P;M)$.

@@

There is one other subtlety introduced by the use of in-sample mean squared error. Technically speaking, elements $f \in H^s(P)$ are equivalence classes, defined only up to a set of measure zero. Thus one cannot speak of the pointwise evaluation $f_0(X_i)$, as we do by defining our target of estimation to be $f_0(X_i)$, $i=1,\ldots,n$, until one selects \emph{representatives}. When $s > d/2$, every element $f$ of $H^s(P)$ admits a continuous version $f^{\ast}$, and it is standard to pick this as the favored representative. When $s \leq d/2$, some elements in $H^s(P)$ do not have any continuous version; however they admit a \emph{quasi-continuous} version \citep{evans15} known as the \emph{precise representative}, and we use this representative. To be clear, however, it does not really matter which representative we choose. Since all versions agree except on a set of measure zero, and since $P$ is absolutely continuous with respect to Lebesgue measure (in Model~\ref{def:model_flat_euclidean}) or the volume form $d\mu$ (in Model~\ref{def:model_manifold}), with probability $1$ any two versions $g_0, h_0 \in f_0$ will satisfy $g_0(X_i) = h_0(X_i)$ for all $i = 1,\ldots,n$. The bottom line is that we can use the notation $f_0(X_i)$ without fear of ambiguity or confusion.

@@

\paragraph{Estimation.} When the graph radius $\varepsilon$ and number of eigenvectors $K$ are chosen appropriately, we show in Theorem~\ref{thm:laplacian_eigenmaps_estimation_fo} that the estimator $\wh{f}$ achieves the minimax rate over $H^1(P;M)$, when error is measured in $\|\cdot\|_{n}$.
\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{0}
	\item 
	\label{asmp:parameters_estimation_fo} 
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy the following inequalities:
	\begin{equation}\\
	\label{eqn:radius_fo} 
	\biggl(\frac{\log n}{n}\biggr)^{1/d} \ll \varepsilon \ll \min\{1,K^{-1/d}\},
	\end{equation}
	and 
	\begin{equation}
	\label{eqn:eigenvector_estimation_fo} 
	K \asymp \min\Bigl\{(M^2n)^{d/(2 + d)} \vee 1, n\Bigr\}
	\end{equation}
\end{enumerate}
\begin{theorem}
	\label{thm:laplacian_eigenmaps_estimation_fo}
	Suppose Model~\ref{def:model_flat_euclidean}, and additionally $f_0 \in H^1(P,M)$. There exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds all $n$ larger than $N$ and for any $\delta \in (0,1)$: if the Laplacian eigenmaps estimator $\wh{f}$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_estimation_fo}, then
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_estimation_fo}
	\|\wh{f} - f_0\|_n^2 \leq C\Bigl(\frac{1}{\delta}M^2(M^2n)^{-2/(2 + d)} \wedge 1\Bigr) \vee \frac{1}{n},
	\end{equation}
	with probability at least $1 - \delta - Cn\exp(-cn\varepsilon^d)$.
\end{theorem}
From~\eqref{eqn:laplacian_eigenmaps_estimation_fo} it follows immediately that with high probability $\|\wh{f} - f_0\|_n^2 \lesssim M^2(M^2n)^{-2/(2 + d)}$ whenever $n^{-1/2} \lesssim M \lesssim n^{1/d}$, and is thus an in-sample minimax rate-optimal estimator.

Some other remarks:
\begin{itemize}
	\item When $M \ll n^{-1/2}$ or $M \gg n^{1/d}$, the rate in~\eqref{eqn:laplacian_eigenmaps_estimation_fo} looks different than~\eqref{eqn:sobolev_space_minimax_estimation_rate}. In particular, when $M \ll n^{-1/2}$, then computing Laplacian eigenmaps with $K = 1$ achieves the parametric rate $\|\wh{f} - f_0\|_n^2 \lesssim n^{-1}$, and the zero-estimator $\wh{f} = 0$ achieves the better rate $\|\wh{f} - f_0\|_n^2 \lesssim M^2$, but neither match the minimax rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. On the other hand, when $M \gg n^{1/d}$, then computing Laplacian eigenmaps with $K = n$ achieves the rate $\|\wh{f} - f_0\|_n^2 \asymp 1$, which is better than the rate in~\eqref{eqn:sobolev_space_minimax_estimation_rate}. However, in truth these are edge cases, which do not fall neatly into the framework of nonparametric regression.
	\item The lower bound on $\varepsilon$ imposed by~\eqref{eqn:radius_fo} is on the order of the connectivity threshold, the smallest length scale at which the resulting graph will still, with high probability, be connected. On the other hand, as we will see in Section~\ref{subsec:analysis}, the upper bound on $\varepsilon$ is needed to ensure that the graph eigenvalue $\lambda_K$ is on the same order as the continuum eigenvalue $\lambda_K(\Delta_P)$. Finally, we choose $K \asymp (M^2n)^{2d/(2 + d)}$ to optimally trade-off bias and variance.
	\item We note that the ranges~\eqref{eqn:radius_fo} and~\eqref{eqn:eigenvector_estimation_fo} depend on unknown quantities such as the dimension $d$ and radius of the Sobolev ball $M$. In practice, one typically tunes hyper-parameters by sample-splitting or cross-validation. However, because the estimator $\wh{f}$ is defined only in-sample, we cannot use such methods to select the graph radius $\varepsilon$, or number of eigenvectors $K$. We return to this issue upon considering out-of-sample extensions of $\wh{f}$ later, in Section~\ref{sec:out_of_sample}.
	\item The upper bound given in Equation~\eqref{eqn:laplacian_eigenmaps_estimation_fo} holds with probability $1 - \delta - Cn\exp(-cn\varepsilon^d)$. Under the stronger assumption that $f_0 \in C^1(\mc{X};M)$ we can replace the factor of $\delta$ by the sharper $\delta^2/n$, which is less than $\delta$ because $\delta \in (0,1)$. 
\end{itemize}

\paragraph{Testing.} Consider the test $\varphi = \1\{\wh{T} \geq t_{a}\}$, where $t_{a}$ is the threshold
\begin{equation*}
t_{a} := \frac{K}{n} + \frac{1}{n}\sqrt{\frac{2K}{a}}.
\end{equation*}
This choice of threshold $t_{a}$ guarantees that $\varphi$ is a level-$a$ test. Moreover, when $\varepsilon$ and $K$ are chosen appropriately, the test $\varphi$ has negligible Type II error against alternatives separated from the null by at least $\|f_0\|_{P}^2 \gtrsim M^2(M^2n)^{-4/(4 + d)}$, whenever $d < 4$.

\begin{enumerate}[label=(A\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:parameters_testing_fo}
	The graph radius $\varepsilon$ and the number of eigenvectors $K$ satisfy~\eqref{eqn:radius_fo}. Additionally,
	\begin{equation}
	\label{eqn:eigenvector_testing_fo}
	K \asymp \min\Bigl\{(M^2n)^{2d/(4 + d)} \vee 1, n\Bigr\}.
	\end{equation}
\end{enumerate}
In the following, let $p(d) = \min\{1,2/d\}$.
\begin{theorem}
	\label{thm:laplacian_eigenmaps_testing_fo}
	Fix $a,b \in (0,1)$. Suppose Model~\ref{def:model_flat_euclidean}. Then $\mathbb{E}_0[\varphi] \leq a$, i.e $\varphi$ is a level-$a$ test. Suppose additionally $f_0 \in H^1(P,M)$, and that $d < 4$. Then there exist constants $c,C$ and $N$ that do not depend on $f_0$, such that the following statement holds for all $n$ larger than $N$: if the Laplacian eigenmaps test $\varphi$ is computed with parameters $\varepsilon$ and $K$ that satisfy~\ref{asmp:parameters_testing_fo}, and if $f_0$ satisfies
	\begin{equation}
	\label{eqn:laplacian_eigenmaps_testing_criticalradius_fo}
	\|f_0\|_P^2 \geq \frac{C}{b}\biggl(\Bigl(M^2(M^2n)^{-4/(4 + d) } \wedge n^{-1/2}\Bigr)\biggl[\sqrt{\frac{1}{a}} + \frac{1}{b}\biggr] \vee \frac{M^2}{b n^{p(d)}} \biggr) \vee \frac{1}{n}
	\end{equation}
	then $\Ebb_{f_0}[1 - \phi] \leq b$.
\end{theorem}
Although~\eqref{eqn:laplacian_eigenmaps_testing_criticalradius_fo} involves taking the maximum of several different terms, the important takeaway of Theorem~\ref{thm:laplacian_eigenmaps_testing_fo} is that if $n^{-1/2} \lesssim M^2 \ll n^{1/4}$ ($d = 1$) or $n^{-1/2} \lesssim M^2 \ll n^{2/d - 1/2}$ ($d = 2$ or $d = 3$), then $\varphi$ is a minimax rate-optimal test over $H^1(P,M)$.
